---
title: "GOM-Rpath"
author: "Sarah Weisberg"
date: "10/1/2020"
bibliography: biblio.json
output:
  html_document: default
  pdf_document: default
  word_document: default
---


```{r setup, echo=FALSE}
library(knitr)
knitr::opts_chunk$set(
  comment = '#>',
  collapse = T)
library(Rpath); library(data.table); library(Survdat); library(ggplot2); library(rgdal); library(here)

#install.packages("remotes")
#remotes::install_github("NOAA-EDAB/Rpath")
#devtools::install_github('slucey/RSurvey/Survdat')

```

## Introduction
This repository contains all of the background code for a mass-balanced food web model of the Gulf of Maine (GOM), which is built using the open-sourced mass balance algorithms of Rpath [@lucey2020]. The GOM model has been parameterized for the 1980-1985 time period, and fit to time series between 1985-2020. 52 functional groups are represented in the model, including 31 single fish species of commercial and/or ecological importance in the GOM. 

Biomass estimates are derived primarily from NOAA's bottom trawl survey. For species and groups not well-covered by the survey -- such as lower trophic levels and marine mammals -- estimates are from the Energy Modeling and Analysis eXercise (EMAX) completed by the Northeast Fisheries Science Center (NEFSC) [@link2006]. The latter was a large-scale effort among dozens of researches to collate known biological parameter estimates for key species and groups across the Northeast US Continental Shelf Ecosystem and produced mass-balanced models of the GOM, Georges Bank (GB), Southern New England and the Mid-Atlantic Bight (MAB) -- efforts that have been foundational towards holistic understanding of the region. However, the EMAX models are highly aggregated, with 31 functional groups; more recent models have attempted to increase resolution and thereby allow for more nuanced questions to be asked and answered. The present model aims to do the same. Additionally, it is created alongside updated Rpath models of both GB and the MAB, in order to facilitate comparative analyses. In keeping with the GB fall model, biomass estimates are based on fall survey data. 
As in the GB model, initial biological parameter estimates are derived from either the EMAX estimates or the more recent model by Buchheister et al. [-@buchheister]. Where appropriate, estimates were also pulled from Heymans (2001) or Dias et al. [@dias2019].

- key data sources = bottom trawl survey, NOAA landings data (aggregated  stomach content analysis data for diet matrix, EMAX and NWACS models for species not collected in trawl.
- I define GOM as bottom trawl strata 1220, 1240, 1260:1290, 1360:1400, 3560:3830 and statistical areas 500, 510, 512:515, 521 & 522 for landings analysis.
- Code is frequently adapted from Sean Lucey's tools for survey analysis (cite) & Rpath model creating
- Group structure within model designed to be able to be analogous to Georges Bank Rpath model

## Generating list of functional groups
-add more detail on how these groups were derived
-Cut-off of biomass from trawl data to be considered as single species
-Included if the biomass exceeded that threshold at any time during the time series
-Groups in EMAX but not in trawl are included as well
-Macrobenthos estimates are from EMAX?


```{r groups, echo=TRUE}
GOM.groups <-as.data.table(c('Phytoplankton', 'Bacteria', 'Microzooplankton', 'GelZooplankton','LgCopepods', 'SmCopepods', 'Micronekton', 'Illex','Loligo', 'OtherCephalopods','Macrobenthos','AmLobster','OceanPout','SmPelagics','SmFlatfishes','NShrimp','OtherShrimps','RiverHerring','AtlScallop','OtherPelagics','AtlHerring','Mesopelagics','SouthernDemersals','YTFlounder','Cod','Haddock','AmPlaice','AtlMackerel','Cusk','OtherDemersals','HMS','OtherSkates','RedHake','Barndoor Skate','Fourspot','SmoothDogfish','Pollock','Goosefish','SilverHake','WhiteHake','SpinyDogfish','Redfish','LittleSkate','SeaBirds','Windowpane','WinterFlounder','WitchFlounder','BlackSeaBass','AtlHalibut','Butterfish','Sharks','WinterSkate','Pinnipeds','BaleenWhales','Odontocetes', 'Megabenthos'))
setnames(GOM.groups,'V1','RPATH')

```


## Compiling Biomass Data

### Generating Biomass Estimates of Surveyed Species
Load in NEFSC bottom trawl data ('survdat') from all time, shape file of survey strata ('strata') and species list for matching survey species codes with RPATH codes ('spp'). Note species of lower abundance / biomass will be aggregated into larger functional groups. Calculate mean stratified biomass for each RPATH group per year, using fall data only. Multiply by total area surveyed per year and divide by area swept per tow to generate swept area biomass. Divide by total survey area to arrive at biomass density estimate (t/km^2). Average over 1980-1985.

```{r survey biomass}
#Load data------------------------------------------------------------------

#Load survey data
load('Survdat.RData')

#Load strata
strata <- readOGR('speciescodesandstrata', 'strata')

#Load species code list
load("speciescodesandstrata/Species_codes.RData")

#Generate area table
strat.area <- getarea(strata, 'STRATA')
setnames(strat.area, 'STRATA', 'STRATUM')

#---------------------------------------------------------------------------

#Subset by season & strata set - choose fall for GOM strata
fall.GOM <- survdat[SEASON == 'FALL' & STRATUM %in% c(1220, 1240, 1260:1290, 1360:1400, 3560:3830), ]

#Calculate total survey area
GOM.strat.area <- strat.area[STRATUM %in% c(1220, 1240, 1260:1290, 1360:1400, 3560:3830), sum(Area)]

#Run stratification prep
setnames(fall.GOM, 'EST_YEAR', 'YEAR')
GOM.prep <- stratprep(fall.GOM, strat.area, strat.col = 'STRATUM', area.col ='Area')

#Merge with RPATH names
#Aggregate groups below 0.05 t/km^2 threshold
spp <- spp[!duplicated(spp$SVSPP),]
spp <- spp[RPATH == 'SummerFlounder', RPATH := 'OtherDemersals',]
spp <- spp[RPATH == 'RedCrab', RPATH := 'Macrobenthos',]
spp <- spp[RPATH == 'Scup', RPATH := 'OtherDemersals',]
spp <- spp[RPATH == 'OffHake', RPATH := 'OtherDemersals',]
spp <- spp[RPATH == 'Bluefish', RPATH := 'OtherPelagics',]
spp <- spp[RPATH == 'AmShad', RPATH := 'SmPelagics',]
GOM.prep<-merge(GOM.prep, spp[,list(SVSPP,COMNAME,RPATH,SCINAME)], by = 'SVSPP')

#Calculate stratified means 
mean.biomass <- stratmean(GOM.prep, group.col = 'SVSPP', strat.col = 'STRATUM')

#Merge stratified means with RPATH names
mean.biomass<-merge(mean.biomass,spp[,list(SVSPP,RPATH,SCINAME,Fall.q)], by = 'SVSPP')

#Calculate total biomass estimates
total.biomass <- sweptarea(GOM.prep, mean.biomass, strat.col = 'STRATUM', area.col = 'Area')

#Merge total biomass with RPATH names
total.biomass<-merge(total.biomass,spp[,list(SVSPP,RPATH,SCINAME,Fall.q)], by = 'SVSPP')

#Calcuate biomass / area in mt
total.biomass <- total.biomass[, biomass.t_area   := (tot.biomass*.001)/(Fall.q*GOM.strat.area)]

#Sum multispecies groups
#OtherDemersals
OtherDemersals.Biomass<-total.biomass[RPATH == 'OtherDemersals', sum(biomass.t_area), by = YEAR]
setnames(OtherDemersals.Biomass,'V1','Biomass')
OtherDemersals.80s <- OtherDemersals.Biomass[YEAR %in% 1980:1985, mean(Biomass)]

#SouthernDemersals
SouthernDemersals.Biomass<-total.biomass[RPATH == 'SouthernDemersals', sum(biomass.t_area), by = YEAR]
setnames(SouthernDemersals.Biomass,'V1','Biomass')
SouthernDemersals.80s <- SouthernDemersals.Biomass[YEAR %in% 1980:1985, mean(Biomass)]

#SmFlatfishes
SmFlatfishes.Biomass<-total.biomass[RPATH == 'SmFlatfishes', sum(biomass.t_area), by = YEAR]
setnames(SmFlatfishes.Biomass,'V1','Biomass')
SmFlatFishes.80s <- SmFlatfishes.Biomass[YEAR %in% 1980:1985, mean(Biomass)]

#OtherPelagics
OtherPelagics.Biomass<-total.biomass[RPATH == 'OtherPelagics', sum(biomass.t_area), by = YEAR]
setnames(OtherPelagics.Biomass,'V1','Biomass')
OtherPelagics.80s <- OtherPelagics.Biomass[YEAR %in% 1980:1985, mean(Biomass)]

#OtherCephalopods
OtherCeph.Biomass<-total.biomass[RPATH == 'OtherCephalopods', sum(biomass.t_area), by = YEAR]
setnames(OtherCeph.Biomass,'V1','Biomass')
OtherCeph.80s <- OtherCeph.Biomass[YEAR %in% 1980:1985, mean(Biomass)]

#Mesopelagics
Mesopelagics.Biomass<-total.biomass[RPATH == 'Mesopelagics', sum(biomass.t_area), by = YEAR]
setnames(Mesopelagics.Biomass,'V1','Biomass')
Mesopelagics.80s <- Mesopelagics.Biomass[YEAR %in% 1980:1985, mean(Biomass)]

#SmPelagics
SmPelagics.Biomass<-total.biomass[RPATH == 'SmPelagics', sum(biomass.t_area), by = YEAR]
setnames(SmPelagics.Biomass,'V1','Biomass')
SmPelagics.80s <- SmPelagics.Biomass[YEAR %in% 1980:1985, mean(Biomass)]

#OtherShrimps
OtherShrimps <-total.biomass[RPATH == 'OtherShrimps']
OtherShrimps.Biomass<-total.biomass[RPATH == 'OtherShrimps', sum(biomass.t_area), by = YEAR]
setnames(OtherShrimps.Biomass,'V1','Biomass')
OtherShrimps.80s <- OtherShrimps.Biomass[YEAR %in% 1980:1985, mean(Biomass)]

#OtherSkates
OtherSkates.Biomass<-total.biomass[RPATH == 'OtherSkates', sum(biomass.t_area), by = YEAR]
setnames(OtherSkates.Biomass,'V1','Biomass')
OtherSkates.80s <- OtherSkates.Biomass[YEAR %in% 1980:1985, mean(Biomass)]

#Average over 1980-1985
GOM.biomass.80s <- total.biomass[YEAR %in% 1980:1985, mean(biomass.t_area), by = RPATH]
setnames(GOM.biomass.80s,'V1','Biomass')
GOM.biomass.80s<-GOM.biomass.80s[RPATH == 'OtherDemersals', Biomass := OtherDemersals.80s,]
GOM.biomass.80s<-GOM.biomass.80s[RPATH == 'Mesopelagics', Biomass := Mesopelagics.80s,]
GOM.biomass.80s<-GOM.biomass.80s[RPATH == 'OtherCephalopods', Biomass := OtherCeph.80s,]
GOM.biomass.80s<-GOM.biomass.80s[RPATH == 'OtherPelagics', Biomass := OtherPelagics.80s,]
GOM.biomass.80s<-GOM.biomass.80s[RPATH == 'OtherShrimps', Biomass := OtherShrimps.80s,]
GOM.biomass.80s<-GOM.biomass.80s[RPATH == 'OtherSkates', Biomass := OtherSkates.80s,]
GOM.biomass.80s<-GOM.biomass.80s[RPATH == 'SmFlatfishes', Biomass := SmFlatFishes.80s,]
GOM.biomass.80s<-GOM.biomass.80s[RPATH == 'SmPelagics', Biomass := SmPelagics.80s,]
GOM.biomass.80s<-GOM.biomass.80s[RPATH == 'SouthernDemersals', Biomass := SouthernDemersals.80s,]


kable(GOM.biomass.80s, booktabs = F, align = 'c')

#Output results to csv
write.csv(GOM.biomass.80s, 'Mean Biomass_GoM_1980-85.csv')

#Average over 2013-18
GOM.biomass.2010s <- total.biomass[YEAR %in% 2013:2018, mean(biomass.t_area), by = RPATH]
setnames(GOM.biomass.2010s,'V1','Biomass')


#Output results to csv
write.csv(GOM.biomass.2010s, 'Mean Biomass_GoM_2013-18.csv')
```

##Pulling biomass estimates from EMAX model
-consider adding code to pull data from Ecobase
-using EMAX biomass esimates for non-trawled species

```{r EMAX biomass}
#Load data------------------------------------------------------------------

#load EMAX model
GOM.EMAX<-as.data.table(read.csv('GOM_EMAX_params.csv'))

#User created functions-----------------------------------------------------
'%notin%' <-Negate('%in%')
#---------------------------------------------------------------------------

##Merge group list with survey biomass estimates
GOM.groups<-merge(GOM.biomass.80s, GOM.groups, by = 'RPATH', all.y=TRUE)
setnames(GOM.groups,'V1','Biomass', skip_absent = TRUE)

#Filter EMAX model for group, biomass
setnames(GOM.EMAX,'model.Group','RPATH')
setnames(GOM.EMAX,'model.Biomass','Biomass')
GOM.EMAX<-GOM.EMAX[, c('RPATH','Biomass')]

#Change EMAX names to match RPATH names
GOM.EMAX<-GOM.EMAX[RPATH == 'Small copepods', RPATH := 'SmCopepods',]
GOM.EMAX<-GOM.EMAX[RPATH == 'Large Copepods', RPATH := 'LgCopepods',]
GOM.EMAX<-GOM.EMAX[RPATH == 'Sea Birds', RPATH := 'SeaBirds',]
GOM.EMAX<-GOM.EMAX[RPATH %like% 'Macro', RPATH:='Macrobenthos',]
GOM.EMAX<-GOM.EMAX[RPATH %like% 'Baleen', RPATH := 'BaleenWhales',]
GOM.EMAX<-GOM.EMAX[RPATH %like% 'Gel',RPATH := 'GelZooplankton',]
GOM.EMAX<-GOM.EMAX[RPATH %like% 'Phyto', RPATH :='Phytoplankton',]
GOM.EMAX<-GOM.EMAX[RPATH %like% 'Megabenthos',RPATH := 'Mega',]
GOM.EMAX<-GOM.EMAX[RPATH %like% 'Pelagics',RPATH := 'Pelagics',]
GOM.EMAX<-GOM.EMAX[RPATH %like% 'Demersals',RPATH := 'Demersals',]
GOM.EMAX<-GOM.EMAX[RPATH %like% 'Sharks', RPATH :='Sharks']

#Aggregate EMAX macrobenthos groups into one 
BenthosBiomass<-GOM.EMAX[RPATH == 'Macrobenthos', sum(Biomass),]
GOM.EMAX<-GOM.EMAX[RPATH == 'Macrobenthos', Biomass:=BenthosBiomass,]

#Remove groups not included in Rpath model
GOM.EMAX<- GOM.EMAX[RPATH %notin% c('Larval-juv fish- all','Shrimp et al.','Mega','Pelagics','Demersals','Discard','Detritus-POC','Fishery'),]
GOM.EMAX<-unique(GOM.EMAX)

#Merge survey and EMAX biomass estimates
GOM.biomass.80s<-merge(GOM.groups,GOM.EMAX,by='RPATH', all=TRUE)
GOM.biomass.80s<-GOM.biomass.80s[is.na(Biomass.y) , Biomass.y := Biomass.x]
GOM.biomass.80s<-GOM.biomass.80s[,Biomass.x :=NULL]
setnames(GOM.biomass.80s,"Biomass.y","Biomass")

#save as .csv, generate table
write.csv(GOM.biomass.80s, 'GOM_Biomass.csv')

kable(GOM.biomass.80s, booktabs = T, align = 'c')

```

#Generating fleet strucutre
-From commercial landings data
-Use statistical areas 500, 510, 512:515, 521, 522
-Segmented by major gear type
-Esimated mean landings for 1980-85, divide by total area

```{r fleets, echo=TRUE}

#Load data------------------------------------------------------------------
#load commercial landings data 
load("~/Desktop/RPath-GOM-workshop/comland_meatwt_deflated_stat_areas.RData")

#Load shape files and calculate total area
stat.areas <- readOGR('gis','Statistical_Areas_2010')

#Calculate total commercial fishing area included in GOM model
stat.areas.area <- getarea(stat.areas, 'Id')
GOM.com.area <- stat.areas.area[Id %in% c(500, 510, 512:515, 521, 522), sum(Area)]

#delete NAs
spp <- unique(spp[!is.na(NESPP3), list(NESPP3, RPATH)])
comland <- comland[!is.na(NESPP3)]

#check for duplicate NESPP3 entries, reassign, remove duplicates
spp.dupes<-spp[duplicated(NESPP3)]
spp<-spp[NESPP3 %in% spp.dupes$NESPP3, RPATH := 'OtherDemersals']
spp <- unique(spp)

#merge landings with species codes
landings <- as.data.table(merge(comland, spp, by = 'NESPP3'))

#subset for GOM 
GOM.stat.areas <- c(500, 510, 512:515, 521, 522)
GOM.landings <- landings[AREA %in% GOM.stat.areas, ]

#Decouple small and large otter trawl fisheries
GOM.landings[SIZE == 'small' & GEAR == 'otter', GEAR := 'otter.sm']
GOM.landings[SIZE == 'large' & GEAR == 'otter', GEAR := 'otter.lg']

#Sum by Rpath group
GOM.sums <- GOM.landings[, sum(SPPLIVMT), by = c('YEAR', 'RPATH', 'GEAR')]
setnames(GOM.sums, 'V1', 'SPPLIVMT')

#Subset & sum for 1980-1985
GOM.landings <- GOM.landings[YEAR %in% 1980:1985]
GOM.sums <- GOM.landings[, sum(SPPLIVMT), by = c('YEAR', 'RPATH', 'GEAR')]
setnames(GOM.sums, 'V1', 'SPPLIVMT')

#Get five year means
GOM.landings <- GOM.sums[YEAR %in% 1980:1985, mean(SPPLIVMT), by = c('RPATH', 'GEAR')]
setnames(GOM.landings, 'V1', 'SPPLIVMT')

#Calculate landings in mt / km^2
GOM.landings <- GOM.landings[, SPPLIVMT := SPPLIVMT / GOM.com.area]

```

#Estimating biological parameters
-Explain data sources - NWACS, EMAX, GB model
-Add code for pulling data

```{r NWACS}
NWACS <- data.table(Group = c('Phytoplankton', 'Other primary producers', 'Bacteria','Microzooplankton', 'Small Copepods', 'Large Copepods','Gelatinous zooplankton', 'Micronekton', 'Macrobenthos- polychaetes', 'Macrobenthos- crustaceans', 'Macrobenthos- molluscs', 'Macrobenthos- other','Megabenthos- filterers', 'Megabenthos- other', 'Shrimp and Similar Species', 'Mesopelagics','Atlantic herring', 'Alosines', 'Atlantic menhaden (S)', 'Atlantic menhaden (M)', 'Atlantic menhaden (L)', 'Anchovies', 'Atlantic mackerel', 'Squid', 'Butterfish','Small pelagics- other', 'Bluefish (S)', 'Bluefish (M)', 'Bluefish (L)', 'Striped bass (S)', 'Striped bass (M)','Striped bass (L)', 'Weakfish (S)', 'Weakfish (M)', 'Weakfish (L)', 'Spiny dogfish (S)', 'Spiny dogfish (L)','Atlantic cod (S)', 'Atlantic cod (M)', 'Atlantic Cod (L)','Haddock', 'Hake', 'Atlantic croaker','Yellowtail flounder (S)', 'Yellowtail flounder (L)', 'Summer flounder (S)', 'Summer flounder (L)', 'Skates','Demersal benthivores - other', 'Demersal piscivores - other','Demersal omnivores - other', 'Medium pelagics - other','Sharks - coastal', 'Sharks - pelagic', 'Large pelagics (HMS)', 'Pinnipeds', 'Baleen whales', 'Odontocetes', 'Seabirds', 'Nearshore pisc birds', 'Detritus'),Biomass = c(30.000, 1.605, 7.700, 7.000, 16.000, 17.966, 6.349, 7.654, 17.452, 7.000, 8.340, 21.000, 5.500, 4.498,  0.470, 0.090, 1.650, 0.200, 0.371, 2.048, 1.200,1.100, 1.740, 1.267, 1.488, 1.400, 0.017, 0.160, 0.509, 0.003, 0.022, 0.022, 0.006, 0.038, 0.036,0.337, 0.800, 0.055, 0.144, 0.277, 0.254, 1.100,0.350, 0.007, 0.187, 0.010, 0.159, 1.000, 2.300,1.300, 1.100, 0.021, 0.008, 0.016, 0.070, 0.035,0.464, 0.060, 0.007, 0.007, 52.6),PB = c(180.700, 55.570, 91.250, 85.000, 46.000, 46.000, 40.000,14.250, 2.500, 3.600, 2.200, 2.000, 1.200, 2.300, 2.000,1.100, 1.100, 1.300, 1.900, 1.309, 0.756, 2.200, 0.550,5.720, 1.312, 1.200, 3.900, 0.900, 0.310, 1.500, 0.526, 0.317, 3.300, 0.900, 1.000, 0.321, 0.321, 1.087, 1.125, 0.700, 0.700, 1.296, 0.994, 2.700, 0.850, 2.200, 1.050,0.250, 0.550, 0.450, 0.550, 0.450, 0.200, 0.113, 0.579, 0.075, 0.040, 0.040, 0.279, 0.279, NA), QB = c(0.000, 0.000, 380.208, 283.400, 140.000, 150.000, 145.326,85.497, 17.500, 21.000, 13.949, 16.059, 6.660, 15.533,6.660, 3.700, 3.700, 4.400, 15.860, 6.643, 3.785, 7.333, 2.170, 19.000, 4.230, 4.000, 20.935, 6.093, 3.139, 10.265, 3.429, 1.820, 13.520, 4.689, 2.803, 3.519, 1.810, 5.059, 2.603, 1.500, 3.000, 3.850, 3.550, 12.168, 2.900, 10.283, 2.900, 0.900, 1.833, 1.500, 1.833, 1.838, 1.247,0.690, 6.794, 5.581, 3.217, 14.301, 80.000, 80.000, NA), RPATH = c('Phytoplankton', 'Phytoplankton', 'Bacteria', 'Microzooplankton', 'Mesozooplankton', 'Mesozooplankton','GelZooplankton', 'Micronekton',rep('Macrobenthos', 4), 'AtlScallop', 'AmLobster','OtherShrimps', 'Mesopelagics', 'AtlHerring', 'SmPelagics',rep('OtherPelagics', 3), 'SmPelagics', 'AtlMackerel','Illex', 'Butterfish', 'SmPelagics', rep('Bluefish', 3),rep('OtherPelagics', 3), rep('SouthernDemersals', 3), rep('SpinyDogfish', 2), rep('Cod', 3), 'Haddock','RedHake', 'SouthernDemersals', rep('YTFlounder', 2),rep('SummerFlounder', 2), 'WinterSkate', rep('OtherDemersals', 3), 'OtherPelagics', rep('Sharks', 2),'HMS', 'Seals', 'BalWhale', 'ToothWhale', 'Seabirds','Seabirds', 'Detritus'))

#Calculate weighted mean
NWACS[, Rpath.bio := sum(Biomass), by = RPATH]
NWACS[, PB.weight := Biomass * PB]
NWACS[, new.PB := sum(PB.weight) / Rpath.bio, by = RPATH]
NWACS[, QB.weight := Biomass * QB]
NWACS[, new.QB := sum(QB.weight) / Rpath.bio, by = RPATH]
```

```{r biomass accumluation}
#RPATH GOM groups
RPATH.GOM <- unique(total.biomass$RPATH, na.rm=true)

#Write a loop to run lm for all RPATH groups, extract slope
ba<-c()
p<-c()
for (i in 2:59){
    lm<-lm(biomass.t_area~YEAR, data = subset(total.biomass, RPATH == RPATH.GOM[i]))
    spF <- as.numeric(summary(lm)$fstatistic)
    p[i] <- pf(spF[1], spF[2], spF[3], lower = F)
    ba[i] <-as.numeric(lm[[1]][2])
}

ba<-cbind(RPATH.GOM,ba,p)
```

## References
``` {references}
````
